proof-of-concept_mods

### 2023-10-25 proof-of-concept_mods.Rmd
fixing gam output df

## Significance Testing

scp'd models from cubic to parse

```{r}
read_path <- "/Users/megardn/Desktop/BGD_Repos/combat_biovar/ukb_basic/" #path to bfp models
```

```{r}
#iterate across gamlss objects
model.files <- list.files(path = read_path, pattern = "mod.rds", full.names = TRUE)

sigma.df <- data.frame("file_path" = model.files)
sigma.df <- sigma.df %>%
  mutate(region = sub("_mod\\.rds$", "", basename(as.character(file_path))))

#test significance of sex term in sigma
sigma.df$sig.sex_drop1.p <- lapply(sigma.df$file_path, get.and.drop1.p, moment="sigma", term="sex")

#extract values of sex term from summary table
summary.list <- lapply(sigma.df$file_path, get.summary)
summary.df <- bind_rows(summary.list)
sigma.sex.df <- summary.df %>%
  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
  mutate(pheno_cat = as.factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA)),
    sig = case_when(p.value < 0.05 ~ TRUE,
                    p.value >= 0.05 ~ FALSE),
    sig_bf.corr = case_when(pheno %in% vol_list_global & p.value < (0.05/length(vol_list_global)) ~ TRUE,
                            pheno %in% vol_list_global & p.value >= (0.05/length(vol_list_global)) ~ FALSE,
                            !(pheno %in% vol_list_global) & p.value < (0.05/length(ct_list)) ~ TRUE,
                            !(pheno %in% vol_list_global) & p.value >= (0.05/length(ct_list)) ~ FALSE,
                            TRUE ~ NA),
    label = sub("_[^_]*_", "_", pheno)) #for plotting

#add more info from parcellations
dk.parc <- dk$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

sigma.sex.df2 <- left_join(sigma.sex.df, dk.parc, by="label")
```

@@

```{r}
table(sigma.sex.df2$sig)
table(sigma.sex.df2$sig_bf.corr)
```

```{r, fig.width=11}
sigma.sex.df2 %>%
  dplyr::filter(sig_bf.corr == TRUE) %>%
  ggplot() +
  geom_col(aes(x=reorder(pheno,t_stat), y=t_stat, fill=pheno_cat)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")

sigma.sex.df2 %>%
  dplyr::filter(sig_bf.corr == TRUE) %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=pheno_cat)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")

sigma.sex.df2 %>%
  dplyr::filter(sig_bf.corr == TRUE) %>%
  ggplot() +
  geom_col(aes(x=region, y=t_stat, fill=hemi)) + 
  facet_wrap(~pheno_cat) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
```

@@

```{r}
sigma.sex.df2 %>%
  dplyr::filter(sig_bf.corr == TRUE) %>%
  group_by(pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_wrap(~pheno_cat) +
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")
```

@@

>>>>

## Review Significance testing

scp'd csvs from cubic

```{r}
drop1.df <- fread("ukb_basic/drop1_tests.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1)
sex.df <- fread("ukb_basic/sex_summary.csv", stringsAsFactors = TRUE)%>%
  dplyr::select(-V1)
sum.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE)%>%
  dplyr::select(-V1)
```

```{r}
#add more info from parcellations
dk.parc <- dk$data %>%
  as.data.frame() %>%
  na.omit() %>%
  dplyr::select(c(hemi, region, label)) %>%
  distinct()

sex.df <- left_join(sex.df, dk.parc, by="label")
sum.df <- left_join(sum.df, dk.parc, by="label")
```

Does the number of regions w significant age effects on mu change depending on combat version?

```{r}
drop1.df %>%
  dplyr::filter(Term=="poly(age_days, 3)" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())
```

No.

Does the number of regions w significant sex effects on mu change depending on combat version?

```{r}
drop1.df %>%
  dplyr::filter(Term=="sexMale" & Moment == "mu") %>%
  group_by(Dataset) %>%
  summarize(n_sig_regions = sum(sig.drop_bf.corr),
            n_regions = n())

sum.df %>%
  dplyr::filter(term=="sexMale" & parameter == "mu") %>%
  group_by(dataset) %>%
  summarize(n_sig_regions = sum(sig.sum_bf.corr),
            n_regions = n())
```

Slightly. Only combat w/o covariates or lm combat without a ref site are noticeably off. Consistent across drop1 and summary outputs

@@

Non-combatted results

```{r, fig.width=11}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset ==	"ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=reorder(pheno,t_stat), y=t_stat, fill=pheno_cat)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=label, y=t_stat, fill=pheno_cat)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")

sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
  ggplot() +
  geom_col(aes(x=region, y=t_stat, fill=hemi)) + 
  facet_wrap(~pheno_cat) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Significant effects of male sex on sigma across phenotype-specific brain charts")
```

@@

```{r}
sex.df %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & parameter == "sigma" & dataset == "ukb-CN-data-simsites") %>%
  group_by(pheno_cat) %>%
  ggplot() +
  geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = t_stat)) +
  scale_fill_gradient2() +
  facet_wrap(~pheno_cat) +
  theme_void() +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  ggtitle("t-statistic for effect of male sex on sigma")
```

@@

```{r}
# Your input string
input_string <- "xxxx ref.batch = yyyy"

# Check if the input string contains "ref.batch =" using a regular expression
if (grepl("ref\\.batch\\s*=\\s*", input_string)) {
  # Your code for splitting the string
  # Split the string into two parts
  split_string <- unlist(strsplit(input_string, "ref.batch", fixed = TRUE))

  # The split_string will contain two elements
  part1 <- split_string[1]
  part2 <- gsub("=", "", split_string[2])

  # Trim any leading/trailing white spaces in part2
  part2 <- trimws(part2)

  # Output the two parts
  print(part1)
  print(part2)
} else {
  cat("The input string does not contain 'ref.batch ='\n")
}
```
